Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2015-09-02 3dpas_intro.tex
checkpoint.

\begin{abstract}
%   Many problems at the forefront of science, engineering, medicine,
%   and the social sciences are increasingly complex and
%   interdisciplinary due to the plethora of data sources and
%   computational methods available today.  A common feature across many
%   of these problems is the amount and diversity of data and
%   computation that must be integrated to yield insights. Such data is
%   increasingly large-scale and distributed: the datasets, their
%   locations, and associated scheduling decisions are
%   time-dependent.

@@

The landscape of scientific and enterprise
computing is being fundamentally altered by the current surge in data. The
resulting upheaval, grandiosely named BigData science by the Lilliputians
in the data science community, represents a new reality where
the volumes of data produced by scientific applications are increasing
rapidly, driven by advancing technologies (e.g., increasing compute
capacity and higher resolution sensors) and decreasing costs for data
acquisition and storage~\cite{tony-propaganda}. With the increase in data
volumes, types, and sources of data, an important concomitant
development has been an increase in the requirement for and complexity
in the processing of data. Data are often naturally distributed and have
to be processed in situ, and occasionally data that are localized
have to be distributed in order to be processed, either for reasons
of performance or for collaborative analytics.  Furthermore when
processing large volumes of data, data source/destination, data
production and transfer rates, and data processing capacity can
all vary over time. Thus properties such as resource scheduling, data
placement, and transfer decisions, that were statically determinable at
small scales, emerge as dynamic and important considerations at
large scales of data and distribution.  The BigData science problem
is therefore only partly about large volumes of data.

Our focus is on a subspace of the BigData problem space, namely that
which is characterized by the distributed and dynamic aspects of
data-intensive science. We introduce the concept of Dynamic
Distributed Data-intensive science and refer to this subspace of
BigData science as D3 science.  In order to appreciate the D3 subspace
we will briefly discuss the importance and role of distribution and
the pervasiveness of dynamical aspects.

@@

The challenges of distributed data-intensive applications span beyond
the management and curation of large volumes of data. A critical
requirement is to make sense of this data, i.e., transform the data
into information en route to assembling knowledge. Among other things,
the transformation from data to information requires extensive amounts
of computational processing of the data; thus the data-intensive
nature of the BigData science problem is strongly coupled with
compute-intensive requirements.  As the data are often distributed, a
consequence is that many attributes of distributed systems permeate to
data. Thus almost naturally, distributed data have attributes of
dynamism, both at the data level as well as the infrastructure
level.

That is not to say all data-intensive applications have elements of
both data and compute dynamism and are always distributed, but it is a
reflection that at extreme scales, dynamism and and distribution
become increasingly important and often correlated issues.  Thus,
analyzing the applications along an analytical space spanned by the
three ``D''s of data-intensive, distributed, and dynamic provides a
general and effective framework.  The diverse set of requirements to
support D3 scenarios points to the need for a careful understanding of
D3 characteristics as well as sophisticated, extensible, and
skillfully architected infrastructure and programming systems.

>>>>

\begin{abstract}

%   Many problems at the forefront of science, engineering, medicine,
%   and the social sciences are increasingly complex and
%   interdisciplinary due to the plethora of data sources and
%   computational methods available today.  A common feature across many
%   of these problems is the amount and diversity of data and
%   computation that must be integrated to yield insights. Such data is
%   increasingly large-scale and distributed: the datasets, their
%   locations, and associated scheduling decisions are
%   time-dependent.

@@

The landscape of scientific and enterprise computing is being fundamentally
altered by the current surge in data. The resulting upheaval represents a new
reality where the volumes of data produced by scientific applications are
increasing rapidly, driven by advancing technologies (e.g., increasing compute
capacity and higher resolution sensors) and decreasing costs for data
acquisition and storage~\cite{tony-propaganda}. With the increase in data
volumes, types, and sources of data, an important concomitant development has
been an increase in the requirement for and complexity in the processing of
data. Data are often naturally distributed and have to be processed in situ, and
occasionally data that are localized have to be distributed in order to be
processed, either for reasons of performance or for collaborative analytics.
Furthermore when processing large volumes of data, data source/destination, data
production and transfer rates, and data processing capacity can all vary over
time.

The set of challenges is only partly defined by the large volumes of
data. Properties such as resource scheduling, data placement, and transfer
decisions, that were statically determinable at small scales, emerge as dynamic
and important considerations at large scales of data and distribution.

Our focus is on a subspace of the large set of problems associated with
data-intensive sciences, namely that which is characterized by the distributed
and dynamic aspects. We introduce the concept of Dynamic Distributed
Data-intensive science and refer to this subspace of data-intensive science as
D3 science.  In order to appreciate the D3 subspace we will briefly discuss the
importance and role of distribution and the pervasiveness of dynamical aspects.

@@

The challenges of distributed data-intensive applications span beyond the
management and curation of large volumes of data. A critical requirement is to
make sense of this data, i.e., transform the data into information en route to
assembling knowledge. Among other things, the transformation from data to
information requires extensive amounts of computational processing of the data;
thus the data-intensive nature of the data-intensive science problem is strongly
coupled with compute-intensive requirements.  As the data are often distributed,
a consequence is that many attributes of distributed systems permeate to
data.

% Thus almost naturally, distributed data have attributes of dynamism, both
% at the data level as well as the infrastructure level.

% That is not to say all data-intensive applications have elements of
% both data and compute dynamism and are always distributed, but it is a
% reflection that at extreme scales, dynamism and and distribution
% become increasingly important and often correlated issues.  Thus,

Analyzing the applications along an analytical space spanned by the
three ``D''s of data-intensive, distributed, and dynamic provides a
general and effective framework.  The diverse set of requirements to
support D3 scenarios points to the need for a careful understanding of
D3 characteristics as well as sophisticated, extensible, and
skillfully architected infrastructure and programming systems.
